---
title: "Tour De France Velogames Daily Update"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# Set stage number
stage_no <- 1
# Step 1: Load league details into R
setwd("/home/a_friend/data_analysis/projects/velogames/2018 - TDF/")
# setwd("C:/aa Simon Lyons/2.0 Work/4.0 Data Analysis/4.6 Projects/velogames/2017 - Vuelta/")
# setwd("C:/aa Simon Lyons/2.0 Work/4.0 Data Analysis/4.6 Projects/velogames/")
league_codes <- read.csv("2018_TDF_velogames_leagues.csv", stringsAsFactors = FALSE)
# Change the number below maually from 1 to 3 to generate data
# for each of the four velogames leagues
league_no <-1
league_code <- league_codes$league_no[league_no]
league <- league_codes$league[league_no]
```

### Race Results - Stage `r stage_no`

<br>
Here's the non-spoiler 4min highlight clip on Youtube:
<br>
<br>
https://youtu.be/GGIb2WT9lkE
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
In the final test for GC contenders it was El Pistolero who both fired the shots and collected the bounty on the villainous slopes of the Angliru. The greatest rider of his generation was finally able to orchestrate a deserving bookend to his career in his final race with a stage victory. In wet conditions Alberto took off at the start of the final climb with his team mate Pantano, reeling in all the breakaway to cross the finish line victoriously in first position. He was followed only 20 seconds later by the hard charging SKY duo of Wout Poels and Chris Froome, effectively securing for Froome the overall win and the amazing Vuelta/TDF double. VIncenzo Nibali lost time but still holds on to third place overall.
<br>
<br>
Results on cycling news:
<br>
<br>
`r paste("http://www.cyclingnews.com/tour-de-france/stage-", stage_no, "/results/", sep = "") `
<br>  
<br>
<br>
     
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

require(XML)
require(knitr)
require(dplyr)
require(ggplot2)
require(rvest)

TDF_table_master <- read.csv(paste("TDF_table_master_", league_code, ".csv", sep = ""))
colnames(TDF_table_master) <- gsub("\n", "", colnames(TDF_table_master))
colnames(TDF_table_master) <- gsub("\\.", "", colnames(TDF_table_master))

# View(TDF_table_master)
# Extract stage results from Cycling News
stage_url <- paste("http://www.cyclingnews.com/races/vuelta-a-espana/stage-", stage_no, "/results/", sep = "")
download.file(stage_url, "stage_url.xml")
stage_html <- htmlParse("stage_url.xml")

# Extract table names/captions
table_captions <- c("Stage Result", unlist(xpathApply(stage_html, "//table/caption", xmlValue)))
# Find location (number) of tables we need for summary (stage result, GC, stage points and overall points)
gc_table_location <- grep("^General Classification", table_captions, fixed = FALSE, ignore.case = TRUE)
points_table_location <- grep("^Points", table_captions, fixed = FALSE, ignore.case = TRUE)
target_tables <- c(1, gc_table_location, points_table_location)

for(tbl in target_tables){
  print(kable(" "))
  print(kable(table_captions[tbl]))
  print(kable(" "))
  stage_table <- as.data.frame(readHTMLTable(stage_html)[tbl])
  colnames(stage_table) <-  as.factor(unlist(stage_table[1,]))
  stage_table <- stage_table[-1, ]
  stage_table <- stage_table[1:10, 1:3]
  print(kable(stage_table), row.names = FALSE)
}
```
   
    
    
### Velogames Results - `r league` - Stage `r stage_no`

  
#### Highest Scoring Riders 
##### for the stage, from our teams
   
The riders earning maximum points for Stage `r stage_no`, and the teams that picked them were:
<br>
<br>
<br>
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring riders for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
rider_stage_score <- TDF_table_master %>% 
  filter(stage == stage_no) %>% 
  select(rider, stage, "Total" = 12) %>% 
  distinct %>% 
  arrange(desc(Total))
# Display as neat table using 'kable' function
# No longer display this table, as I'm using the new table calculated below.
# kable(rider_stage_score[1:10, ])

```
<br>

``` {r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Next add a column with the names of teams against the high scoring stage riders:
for(r in 1:length(rider_stage_score$rider)){
  rider_stage_score[r, 4] <- TDF_table_master %>% 
    filter(rider == rider_stage_score$rider[r]) %>% 
    distinct(team) %>% 
    unlist() %>% 
    paste(collapse = ",")
}
# Give the new column a name
colnames(rider_stage_score)[4] <- "Selected by"

# Display the resulting table
kable(rider_stage_score[1:5, ])

```

<br>
Another take on rider data is provided in the chart below displaying Rider Cost vs the their Total Points accrued so far. The size of each circle is proportional the the number of teams that have selected that rider. 
<br>
<br>
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# Top scoring riders cumulative

# Build a table with unique riders, times selected, total points and cost
  
  # First build a table with individual rider totals
  new_table <- TDF_table_master %>% 
    select(rider, Team,  stage, Cost, "Total" = 12) %>% 
    distinct() %>% 
    group_by(rider, Team, Cost) %>% 
    summarise("Total" = sum(Total)) %>% 
    arrange(rider)   # Arrange by rider so order is matched to next table
  # View(new_table)
  
  
  # Next build table with the number of times each rider has been selected
  rider_times_selected <- TDF_table_master %>% 
    filter(stage == 1) %>% 
    group_by("Rider" = rider) %>% 
    summarise("Times Selected" = n()) %>% 
    arrange(Rider)   # Again arrange by rider to match order from previous table
  # View(rider_times_selected)
  
  # Add times selected column from 2nd table to 1st table
  new_table[ , 5] <- rider_times_selected$`Times Selected`
  colnames(new_table)[5] <- "Times_Selected"
  
  # Create plot showing Total Points, Cost and Times Selected for each rider
  new_plot <- ggplot(data = new_table, aes(x = Cost, y = Total)) +
    geom_point(aes(size = Times_Selected, colour = "blue")) + 
    geom_text(data = subset(new_table, (Total > 100) | (Times_Selected >2)) ,   aes(label=rider),hjust=0, vjust=0, size = 4) + 
    theme(legend.position="bottom") + scale_size(range = c(2,12)) + 
    labs(title = "Rider Points vs Cost",   y = "Total Rider Points") 
  
  new_plot


```
   
  
### Top scoring teams for stage

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring teams for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
team_stage_score <- TDF_table_master %>% 
  filter(stage == stage_no) %>% 
  select(team, directeur, stage, "Total" = 12) %>% 
  group_by(team, directeur) %>% 
  summarise("Stage_Points" = sum(Total)) %>% 
  arrange(desc(Stage_Points))

```
   
<br>  
So `r team_stage_score[1,2]` cleaned up stage `r stage_no`.
<br>  
<br>
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# Display as neat table using 'kable' function
kable(team_stage_score)
# Team scores by stage
team_scores <- TDF_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  arrange(stage, desc(`sum(Tot)`))
ggplot(data = team_scores, aes(x = stage, y = `sum(Tot)`)) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Stage Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

```
   
<br>  
The picture for overall after `r stage_no` stages is:
<br>
<br>
<br>
<br>
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

### Top scoring teams cumulative

# Cumulative team scores by stage
team_cu_scores <- TDF_table_master %>% 
  group_by(stage, team, directeur) %>% 
  summarise(sum(Tot)) %>% 
  group_by(team) %>%
  mutate(cu_stage = cumsum(`sum(Tot)`))
# View(team_cu_scores)
# View(TDF_table_master)

# Plot cumulative team scores
ggplot(data = team_cu_scores, aes(x = stage, y = cu_stage, order = (team_cu_scores$team))) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Cumulative Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

team_overall <- team_cu_scores %>%
  filter(stage == stage_no) %>% 
  select("Team" = team,  "Directeur" = directeur, "Total" = cu_stage) %>% 
  arrange(desc(Total))
kable(team_overall)
# View(team_overall)

```


#### Inside Look

 
Here's a look inside the teams currently holding top spots on the GC podium. 


  
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dlplyr filter looks at the top 4 scoring riders from
# the three teams that made the podium.

# This bit is crapping itself during the 'knit'. Needs work....
stage_cu_podium <- as.data.frame(team_cu_scores) %>%
  filter(stage == as.integer(stage_no)) %>%
  arrange(desc(cu_stage))
# View(stage_cu_podium)

team_cu_rider_scores <- TDF_table_master %>%
  group_by(team, rider) %>%
  summarise("Rider Total" = sum(Tot)) %>%
  filter(team %in% stage_cu_podium$team[1:3]) %>%
  arrange(team, desc(`Rider Total`)) %>%
  top_n(n = 3, wt = `Rider Total`) %>%
  rename("Rider" = rider, "Team" = team)
# View(team_cu_rider_scores)
kable(team_cu_rider_scores, caption = "Rider Scores - Top Teams")

```

<br>
Next up is Stage `r (stage_no + 1)`.
<br>
`r paste("http://www.cyclingnews.com/vuelta-a-espana/stage-", (stage_no + 1), "/", sep = "")`
<br>

### Team Cross Reference
<br>

Here's the team breakdown for each of squads. 
<br>
<br>

``` {r, echo=FALSE, warning=FALSE, message=FALSE}

  # Build a list of the riders selected by the mini-league and the number
  # of times they have been selected
  rider_list <- TDF_table_master %>% 
    filter(stage == 1) %>% 
    group_by(rider) %>% 
    summarise("Times_Selected" = n()) %>% 
    arrange(desc(Times_Selected))
  
  # Build a list of the directeurs in the mini-league
  directeur_list <- TDF_table_master %>% 
    select(directeur) %>% 
    distinct() %>% 
    arrange(directeur)
  
  # Now create the cross-reference matrix which shows which riders have been selected by which directeurs
  for (d in 1:length(directeur_list$directeur)){   # FOR loop to run through the list of team directeurs
    rider_list[ , d+2] <- NA
    colnames(rider_list)[(d+2)] <- toString(directeur_list$directeur[d])
    
    # Build list of team rider for specific team directeur 'd'
    team_list <- TDF_table_master %>% 
      filter(directeur == directeur_list$directeur[d]) %>% 
      select(rider) %>% 
      distinct()
    
    # Next, for each directeur, run through the complete list of mini-league riders
    # and check to see if they are in 
    for (r in 1:length(rider_list$rider)){   # FOR loop to run through the total list of riders in the mini-league
      if(unlist(rider_list[r, 1]) %in% unlist(team_list)){
        rider_list[r, d+2] <- "Yes"
      } else  # End IF statement checking to see if rider exists in the directeur's team
      {rider_list[r, d+2] <- ""}
    }   # End FOR loop 'r'
  }   # END FOR loop running through all of the directeurs
  


kable(rider_list)

```


### DNF Riders

Probably of interest to many of us at this point is which riders are no longer part of the peloton, either through accident, illness or missing the time-cut!


```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Temporary list of DNF riders

# Wikipedia list of 2017 Vuelta riders
wiki_vuelta_2017 <- "https://en.wikipedia.org/wiki/List_of_teams_and_cyclists_in_the_2017_Vuelta_a_Espa%C3%B1a"
# Scrape web data
download.file(wiki_vuelta_2017, "wiki_vuelta_2017.xml")
wiki_vuelta_2017_parse <- read_html("wiki_vuelta_2017.xml")

# Extract table with riders
riders_vuelta <- wiki_vuelta_2017_parse %>% 
   html_nodes(xpath = "//table") %>% 
   html_table(fill = TRUE)
riders_vuelta <- as.data.frame(riders_vuelta[2])

# Filter DNF riders
riders_DNF <- riders_vuelta %>% 
   filter(!Pos. == "") %>% 
   select(Name, Nationality, Team, Pos.)

for (d in 1: length(riders_DNF$Name)){
  riders_DNF[d , 5] <- TDF_table_master %>% 
    filter(rider == riders_DNF$Name[d]) %>% 
    distinct(directeur) %>% 
    unlist() %>% 
    paste(collapse = ",")
}

colnames(riders_DNF)[5] <- "Selected by"

# Print (kable)
print(kable(riders_DNF), row.names = FALSE)

############################


```


