---
title: "Giro Velogames Daily Update"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# Set stage number
stage_no <- 3
# Set league code. Choose from: 673022392215 (AusBus Rouleurs)[1], 760208202214 (Cunning Stunts)[2], 
#                               760302262113 (4051 Rouleurs)[3], 940305525011 (Brissy Boys)[4]
# Step 1: Load league details into R
setwd("/home/a_friend/data_analysis/projects/velogames/")
league_codes <- read.csv("2017_giro_velogames_leagues.csv")
# Change the number below maually from 1 to 4 to generate data
# for each of the four velogames leagues
league_no <- 4
league_code <- league_codes$league_no[league_no]
league <- league_codes$league[league_no]
```

### Race Results - Stage `r stage_no`



Like your flatulent great-Uncle who enjoys a little too much curry, a bit of strong wind came to the Giro party overnight and made its presence known. Just when it looked like the Greipel show was just getting warmed up, the weather played its part and a small groups starting splitting out with under 10 kilometres to go. A dozen riders broke clear, dominated by the Quickstep squad who worked well to deliver their sprinter Fernando Gaviria to victory. The Columbian was followed by Rüdiger Selig and Giacomo Nizzolo. The main peloton, with the favoured sprinters and GC contenders crossed the line a scant 13 seconds later - but an eternity for Ewan and Greipel who will have been expecting a battle for the win. The gap was also enough for Gaviria to wrest the leader's jersey from Greipel.

In GC news, Rohan Dennis crashed and lost 5 mins and Igor Zakarin's team again had to work feverishly to pull him back up to the main group. It will be interesting to see if these efforts have a longer term impact on his team and their ability to support him in the mountains. 

http://www.cyclingnews.com/giro-ditalia/stage-3/results/
   
     
     
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

require(XML)
require(knitr)
require(dplyr)
require(ggplot2)
setwd("/home/a_friend/data_analysis/projects/velogames/")
giro_table_master <- read.csv(paste("giro_table_master_", league_code, ".csv", sep = ""))

# Extract stage results from Cycling News
stage_url <- paste("http://www.cyclingnews.com/races/giro-ditalia-2017/stage-", stage_no, "/results/", sep = "")
download.file(stage_url, "stage_url.xml")
stage_html <- htmlParse("stage_url.xml")

# Extract table names/captions
table_captions <- c("Stage Result", unlist(xpathApply(stage_html, "//table/caption", xmlValue)))
# Find location (number) of tables we need for summary (stage result, GC, stage points and overall points)
gc_table_location <- grep("^General Classification", table_captions, fixed = FALSE, ignore.case = TRUE)
points_table_location <- grep("^Points", table_captions, fixed = FALSE, ignore.case = TRUE)
target_tables <- c(1, gc_table_location, points_table_location)

for(tbl in target_tables){
  print(kable(" "))
  print(kable(table_captions[tbl]))
  print(kable(" "))
  stage_table <- as.data.frame(readHTMLTable(stage_html)[tbl])
  colnames(stage_table) <-  as.factor(unlist(stage_table[1,]))
  stage_table <- stage_table[-1, ]
  stage_table <- stage_table[1:10, 1:3]
  print(kable(stage_table), row.names = FALSE)
}
```
   
    
    
### Velogames Results - `r league` - Stage `r stage_no`

  
#### Highest Scoring Riders 
##### for the stage, from our teams
   
The riders earning maximum points for Stage `r stage_no`, and the teams that picked them were:
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring riders for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
rider_stage_score <- giro_table_master %>% 
  filter(stage == stage_no) %>% 
  select(rider, stage, "Total" = 12) %>% 
  distinct %>% 
  arrange(desc(Total))
# Display as neat table using 'kable' function
# No longer display this table, as I'm using the new table calculated below.
# kable(rider_stage_score[1:10, ])

```


``` {r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Next add a column with the names of teams against the high scoring stage riders:
for(r in 1:length(rider_stage_score$rider)){
  rider_stage_score[r, 4] <- giro_table_master %>% 
    filter(rider == rider_stage_score$rider[r]) %>% 
    distinct(team) %>% 
    unlist() %>% 
    paste(collapse = ",")
}
# Give the new column a name
colnames(rider_stage_score)[4] <- "Selected by"

# Display the resulting table
kable(rider_stage_score[1:10, ])

```



   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# Top scoring riders cumulative
```
   
   
#### Top scoring teams for stage
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring teams for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
team_stage_score <- giro_table_master %>% 
  filter(stage == stage_no) %>% 
  select(team, directeur, stage, "Total" = 12) %>% 
  group_by(team, directeur) %>% 
  summarise("Stage_Points" = sum(Total)) %>% 
  arrange(desc(Stage_Points))

```
   
So `r team_stage_score[1,2]` cleaned up stage 2, courtesy of his Italain stallion Fernando Gaviria. The points will be welcome for the Itchy Nads squad as they look for any scrap of good news in their battle to escape the bottom of the table.
   
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# Display as neat table using 'kable' function
kable(team_stage_score)
# Team scores by stage
team_scores <- giro_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  arrange(stage, desc(`sum(Tot)`))
ggplot(data = team_scores, aes(x = stage, y = `sum(Tot)`)) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Stage Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

```
   
The picture for overall after `r stage_no` stages is:

   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

### Top scoring teams cumulative

# Cumulative team scores by stage
team_cu_scores <- giro_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  group_by(team) %>%
  mutate(cu_stage = cumsum(`sum(Tot)`))


# Plot cumulative team scores
ggplot(data = team_cu_scores, aes(x = stage, y = cu_stage, order = (team_cu_scores$team))) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Cumulative Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")



```
   
#### Inside Look

Here's a look inside the teams carving out early good form. 
   
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dlplyr filter looks at the top 4 scoring riders from
# the three teams that made the podium.

stage_cu_podium <- team_cu_scores %>% 
  filter(stage == 2) %>% 
  arrange(desc(cu_stage))
# View(stage_cu_podium)

team_cu_rider_scores <- giro_table_master %>% 
  group_by(team, rider) %>% 
  summarise("Rider Total" = sum(Tot)) %>% 
  filter(team %in% stage_cu_podium$team[1:3]) %>% 
  arrange(team, desc(`Rider Total`)) %>% 
  top_n(n = 3, wt = `Rider Total`) %>% 
  rename("Rider" = rider, "Team" = team)
kable(team_cu_rider_scores, caption = "Rider Scores - Top Teams")

```

Here's a more useful version of the table I sent out with the Stage 1 update. As well as identifying which riders have been selected the most times in our Velogames squads, it now includes which teams have each rider. It should be helpful in determining which of your squad are essentially a null contribution given they are matched in one or more other teams.



``` {r, echo=FALSE, warning=FALSE, message=FALSE}

# First create the table determining which riders have been selected the most times,
# in descending order:
rider_cost <- giro_table_master %>% 
  filter(stage == 1) %>% 
  group_by("Rider" = rider) %>% 
  summarise("Times Selected" = n()) %>% 
  arrange(desc(`Times Selected`))

# Next add a column with the names of teams against the most popular riders:
for(r in 1:length(rider_cost$Rider)){
  rider_cost[r, 3] <- giro_table_master %>% 
    filter(rider == rider_cost$Rider[r]) %>% 
    distinct(team) %>% 
    unlist() %>% 
    paste(collapse = ",")
}
# Give the new column a name
colnames(rider_cost)[3] <- "Selected by"

kable(rider_cost[1:12,])

```



The race has a rest day tonight as the riders travel off the island of Sardinia and onto the mainland. The sprinters will be missing in action as we see and early mountain top finish - on the occasion on the slopes of the famed Mount Etna!

http://www.cyclingnews.com/giro-ditalia/stage-4/


