---
title: "Giro Velogames Daily Update"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# Set stage number
stage_no <- 21
# Set league code. Choose from: 673022392215 (AusBus Rouleurs)[1], 760208202214 (Cunning Stunts)[2], 
#                               760302262113 (4051 Rouleurs)[3], 940305525011 (Brissy Boys)[4]
# Step 1: Load league details into R
setwd("C:/aa Simon Lyons/2.0 Work/4.0 Data Analysis/4.6 Projects/velogames/Velogames_League_Tables")
league_codes <- read.csv("2017_giro_velogames_leagues.csv")
# Change the number below maually from 1 to 4 to generate data
# for each of the four velogames leagues
league_no <- 1
league_code <- league_codes$league_no[league_no]
league <- league_codes$league[league_no]
```

### Race Results - Stage `r stage_no`
<br>
<br>
<br>
Many of you no doubt already know the result, but here's the non-spoiler 5min highlight clip on Youtube:
<br>
<br>
https://youtu.be/iBVVAbSllJk
<br>
<br>
<br>
<br>
<br>
<br>
Although Time Trial specialist Tom Dumoulin was pipped by countryman Jos van Emden for victory in the final stage of the Giro D'Italia, the Dutchman was most interested in the finishing time of the little Columbian and wearer of the leaders' Maglia Rosa, the last rider to finish. Big Tom need not be concerned as his time gap over Quintana was more than enough to secure his first Grand Tour win. Duloulin's victory in the 100th edition of the Giro was historic for more than one reason, given it's the first Grand Tour victory for a Dutch rider since Joop Zoetemelk won the 1980 Tour de France.
<br>
<br>
Nairo Quintana limited his time loss to finish second overall with last year's winner Vincenzo Nibali claiming the final spot on the podium.
<br>
<br>
Results on cycling news:
<br>
<br>
`r paste("http://www.cyclingnews.com/giro-ditalia/stage-", stage_no, "/results/", sep = "") `
<br>  

     
     
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

require(XML)
require(knitr)
require(dplyr)
require(ggplot2)
setwd("C:/aa Simon Lyons/2.0 Work/4.0 Data Analysis/4.6 Projects/velogames/Velogames_League_Tables")
giro_table_master <- read.csv(paste("giro_table_master_", league_code, ".csv", sep = ""))
colnames(giro_table_master) <- gsub("\n", "", colnames(giro_table_master))
colnames(giro_table_master) <- gsub("\\.", "", colnames(giro_table_master))

# Extract stage results from Cycling News
stage_url <- paste("http://www.cyclingnews.com/races/giro-ditalia-2017/stage-", stage_no, "/results/", sep = "")
download.file(stage_url, "stage_url.xml")
stage_html <- htmlParse("stage_url.xml")

# Extract table names/captions
table_captions <- c("Stage Result", unlist(xpathApply(stage_html, "//table/caption", xmlValue)))
# Find location (number) of tables we need for summary (stage result, GC, stage points and overall points)
gc_table_location <- grep("^General Classification", table_captions, fixed = FALSE, ignore.case = TRUE)
# points_table_location <- grep("^Points", table_captions, fixed = FALSE, ignore.case = TRUE)
target_tables <- c(1, gc_table_location)

for(tbl in target_tables){
  print(kable(" "))
  print(kable(table_captions[tbl]))
  print(kable(" "))
  stage_table <- as.data.frame(readHTMLTable(stage_html)[tbl])
  colnames(stage_table) <-  as.factor(unlist(stage_table[1,]))
  stage_table <- stage_table[-1, ]
  stage_table <- stage_table[1:10, 1:3]
  print(kable(stage_table), row.names = FALSE)
}
```
   
    
    
### Velogames Results - `r league` - Final
<br>
It's worth pointing out the method used by Velogames for calculating points after the final stage. Velogames awards points in the normal manner for Stage 21, but then awards points using an imaginary Stage 22 for Bonus Points based on rider finishing positions in accordance with the game rules:
<br>
<br>
https://www.velogames.com/giro-ditalia/2017/scores.php
<br>
<br>

#### Highest Scoring Riders 
##### for Stage `r stage_no`, from our teams
   
The riders earning maximum points for Stage `r stage_no`, and the teams that picked them were:
<br>
<br>
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring riders for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
rider_stage_score <- giro_table_master %>% 
  filter(stage == stage_no) %>% 
  select(rider, stage, "Total" = 12) %>% 
  distinct %>% 
  arrange(desc(Total))
# Display as neat table using 'kable' function
# No longer display this table, as I'm using the new table calculated below.
# kable(rider_stage_score[1:10, ])

```
<br>
<br>

``` {r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Next add a column with the names of teams against the high scoring stage riders:
for(r in 1:length(rider_stage_score$rider)){
  rider_stage_score[r, 4] <- giro_table_master %>% 
    filter(rider == rider_stage_score$rider[r]) %>% 
    distinct(team) %>% 
    unlist() %>% 
    paste(collapse = ",")
}
# Give the new column a name
colnames(rider_stage_score)[4] <- "Selected by"

# Display the resulting table
kable(rider_stage_score[1:10, ])

```


<br>

#### Highest Scoring Riders 
##### for Stage `r (stage_no+1)`, from our teams
   
The riders earning maximum points from the Bonus Points on offer (Stage `r (stage_no+1)`), and the teams that picked them were:
<br>
<br>
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring riders for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
rider_stage_score <- giro_table_master %>% 
  filter(stage == (stage_no+1)) %>% 
  select(rider, stage, "Total" = 12) %>% 
  distinct %>% 
  arrange(desc(Total))
# Display as neat table using 'kable' function
# No longer display this table, as I'm using the new table calculated below.
# kable(rider_stage_score[1:10, ])

```
<br>
<br>

``` {r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Next add a column with the names of teams against the high scoring stage riders:
for(r in 1:length(rider_stage_score$rider)){
  rider_stage_score[r, 4] <- giro_table_master %>% 
    filter(rider == rider_stage_score$rider[r]) %>% 
    distinct(team) %>% 
    unlist() %>% 
    paste(collapse = ",")
}
# Give the new column a name
colnames(rider_stage_score)[4] <- "Selected by"

# Display the resulting table
kable(rider_stage_score[1:10, ])

```


<br>
The chart below demonstrates what a difference Tom Dumoulin made in the Velogames comp. Whilst Nario and Vincenzo also earnt a stack of points for their teams, their relative cost made them much poorer value. Bobby Jengels stands out as being excellent value as was Pozzovivo. 

   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# Top scoring riders cumulative

# Build a table with unique riders, times selected, total points and cost
  
  # First build a table with individual rider totals
  new_table <- giro_table_master %>% 
    select(rider, Team,  stage, Cost, "Total" = 12) %>% 
    distinct() %>% 
    group_by(rider, Team, Cost) %>% 
    summarise("Total" = sum(Total)) %>% 
    arrange(rider)   # Arrange by rider so order is matched to next table
  # View(new_table)
  
  
  # Next build table with the number of times each rider has been selected
  rider_times_selected <- giro_table_master %>% 
    filter(stage == 1) %>% 
    group_by("Rider" = rider) %>% 
    summarise("Times Selected" = n()) %>% 
    arrange(Rider)   # Again arrange by rider to match order from previous table
  # View(rider_times_selected)
  
  # Add times selected column from 2nd table to 1st table
  new_table[ , 5] <- rider_times_selected$`Times Selected`
  colnames(new_table)[5] <- "Times_Selected"
  
  # Create plot showing Total Points, Cost and Times Selected for each rider
  new_plot <- ggplot(data = new_table, aes(x = Cost, y = Total)) +
    geom_point(aes(size = Times_Selected, colour = "blue")) + 
    geom_text(data = subset(new_table, (Total > 100) | (Times_Selected >2)) ,   aes(label=rider),hjust=0, vjust=0, size = 4) + 
    theme(legend.position="bottom") + scale_size(range = c(2,12)) + 
    labs(title = "Rider Points vs Cost",   y = "Total Rider Points") 
  
  new_plot


```
   
<br>   
<br>

#### Top scoring teams for stage


Points from the ITT:
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring teams for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
team_stage_score <- giro_table_master %>% 
  filter(stage == stage_no) %>% 
  select(team, directeur, stage, "Total" = 12) %>% 
  group_by(team, directeur) %>% 
  summarise("Stage_Points" = sum(Total)) %>% 
  arrange(desc(Stage_Points))

```
   
<br>  
So `r team_stage_score[1,2]` cleaned up stage `r stage_no`.
<br>  
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# Display as neat table using 'kable' function
kable(team_stage_score)
# Team scores by stage
team_scores <- giro_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  arrange(stage, desc(`sum(Tot)`))
# ggplot(data = team_scores, aes(x = stage, y = `sum(Tot)`)) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Stage Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

```
   
<br>  
<br>

Bonus Points (Stage 22):

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring teams for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
team_stage_score <- giro_table_master %>% 
  filter(stage == (stage_no + 1)) %>% 
  select(team, directeur, stage, "Total" = 12) %>% 
  group_by(team, directeur) %>% 
  summarise("Stage_Points" = sum(Total)) %>% 
  arrange(desc(Stage_Points))

```
   
<br>  
So `r team_stage_score[1,2]` cleaned up stage `r stage_no`.
<br>  
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# Display as neat table using 'kable' function
kable(team_stage_score)
# Team scores by stage
team_scores <- giro_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  arrange(stage, desc(`sum(Tot)`))
ggplot(data = team_scores, aes(x = stage, y = `sum(Tot)`)) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Stage Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

```
 


<br>  
<br>
The final overall results after `r stage_no` stages is:
<br>
<br>
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

### Top scoring teams cumulative

# Cumulative team scores by stage
team_cu_scores <- giro_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  group_by(team) %>%
  mutate(cu_stage = cumsum(`sum(Tot)`))

team_overall <- team_cu_scores %>%
  filter(stage == stage_no) %>% 
  select("Team" = team, "Total" = cu_stage) %>% 
  arrange(desc(Total))
kable(team_overall)
# View(team_overall)


# Plot cumulative team scores
ggplot(data = team_cu_scores, aes(x = stage, y = cu_stage, order = (team_cu_scores$team))) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Cumulative Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")



```
   
<br>
   
#### Inside Look
   
Here's a look inside the teams taking top spots on the Velogames GC podium for this 100th edition of the Giro. 
<br>  
   
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dlplyr filter looks at the top 4 scoring riders from
# the three teams that made the podium.

stage_cu_podium <- team_cu_scores %>% 
  filter(stage == stage_no) %>% 
  arrange(desc(cu_stage))
# View(stage_cu_podium)

# I've just changed this to use 'stage_cu_podium' instead of giro_table_master (11 May 17)
# .... AND now I've just changed it back!
team_cu_rider_scores <- giro_table_master %>% 
  group_by(team, rider) %>% 
  summarise("Rider Total" = sum(Tot)) %>% 
  filter(team %in% stage_cu_podium$team[1:3]) %>% 
  arrange(team, desc(`Rider Total`)) %>% 
  top_n(n = 3, wt = `Rider Total`) %>% 
  rename("Rider" = rider, "Team" = team)
kable(team_cu_rider_scores, caption = "Rider Scores - Top Teams")

```

<br>
<br>

### Team Cross Reference
<br>

Here's a more useful version of the table I sent out with the Stage 1 update. As well as identifying which riders have been selected the most times in our Velogames squads, it now includes which teams have each rider. It should be helpful in determining which of your squad are essentially a null contribution given they are matched in one or more other teams.
<br>


``` {r, echo=FALSE, warning=FALSE, message=FALSE}

  # Build a list of the riders selected by the mini-league and the number
  # of times they have been selected
  rider_list <- giro_table_master %>% 
    filter(stage == 1) %>% 
    group_by(rider) %>% 
    summarise("Times_Selected" = n()) %>% 
    arrange(desc(Times_Selected))
  
  # Build a list of the directeurs in the mini-league
  directeur_list <- giro_table_master %>% 
    select(directeur) %>% 
    distinct() %>% 
    arrange(directeur)
  
  # Now create the cross-reference matrix which shows which riders have been selected by which directeurs
  for (d in 1:length(directeur_list$directeur)){   # FOR loop to run through the list of team directeurs
    rider_list[ , d+2] <- NA
    colnames(rider_list)[(d+2)] <- toString(directeur_list$directeur[d])
    
    # Build list of team rider for specific team directeur 'd'
    team_list <- giro_table_master %>% 
      filter(directeur == directeur_list$directeur[d]) %>% 
      select(rider) %>% 
      distinct()
    
    # Next, for each directeur, run through the complete list of mini-league rideres
    # and check to see if they are in 
    for (r in 1:length(rider_list$rider)){   # FOR loop to run through the total list of riders in the mini-league
      if(unlist(rider_list[r, 1]) %in% unlist(team_list)){
        rider_list[r, d+2] <- "Yes"
      } else  # End IF statement checking to see if rider exists in the directeur's team
      {rider_list[r, d+2] <- ""}
    }   # End FOR loop 'r'
  }   # END FOR loop running through all of the directeurs
  


kable(rider_list)

```






