---
title: "TDF Velogames Daily Update"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# Set stage number
stage_no <- 1
# Step 1: Load league details into R
setwd("/home/a_friend/data_analysis/projects/velogames/")
league_codes <- read.csv("2017_tdf_velogames_leagues.csv", stringsAsFactors = FALSE)
# Change the number below maually from 1 to 3 to generate data
# for each of the four velogames leagues
league_no <- 3
league_code <- league_codes$league_no[league_no]
league <- league_codes$league[league_no]
```

### Race Results - Stage `r stage_no`

<br>
<br>
<br>
Here's the non-spoiler 5min highlight clip on Youtube:
<br>
<br>
https://youtu.be/07e_uXhRRm4
<br>
<br>
<br>
<br>
<br>
<br>
Le Tour is underway. I've been away for the weekend and have had to rush the first update, so the details below are a bit rough. I'll smooth things out over the next couple of days. You may know that the first stage was marked by particularly wet weather resulting in some nasty crashes during the short individual time trail. The highest profile of which was Valverde's. The poor Spaniard crashed into a barrier slashing open his leg and breaking his knee cap, ending his stage and Tour participation. Many of the favourites for this time trial took the safe option, including Richie Porte and finished further down than was anticipated. Chris Froome did the best of the GC favrourites.
<br>
<br>
Results on cycling news:
<br>
<br>
`r paste("http://www.cyclingnews.com/tour-de-france/stage-", stage_no, "/results/", sep = "") `
<br>  

     
     
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

require(XML)
require(knitr)
require(dplyr)
require(ggplot2)
setwd("/home/a_friend/data_analysis/projects/velogames/")
tdf_table_master <- read.csv(paste("tdf_table_master_", league_code, ".csv", sep = ""))
colnames(tdf_table_master) <- gsub("\n", "", colnames(tdf_table_master))
colnames(tdf_table_master) <- gsub("\\.", "", colnames(tdf_table_master))

# View(tdf_table_master)
# Extract stage results from Cycling News
stage_url <- paste("http://www.cyclingnews.com/races/tour-de-france/stage-", stage_no, "/results/", sep = "")
download.file(stage_url, "stage_url.xml")
stage_html <- htmlParse("stage_url.xml")

# Extract table names/captions
table_captions <- c("Stage Result", unlist(xpathApply(stage_html, "//table/caption", xmlValue)))
# Find location (number) of tables we need for summary (stage result, GC, stage points and overall points)
gc_table_location <- grep("^General Classification", table_captions, fixed = FALSE, ignore.case = TRUE)
points_table_location <- grep("^Points", table_captions, fixed = FALSE, ignore.case = TRUE)
target_tables <- c(1, gc_table_location, points_table_location)

for(tbl in target_tables){
  print(kable(" "))
  print(kable(table_captions[tbl]))
  print(kable(" "))
  stage_table <- as.data.frame(readHTMLTable(stage_html)[tbl])
  colnames(stage_table) <-  as.factor(unlist(stage_table[1,]))
  stage_table <- stage_table[-1, ]
  stage_table <- stage_table[1:10, 1:3]
  print(kable(stage_table), row.names = FALSE)
}
```
   
    
    
### Velogames Results - `r league` - Stage `r stage_no`

  
#### Highest Scoring Riders 
##### for the stage, from our teams
   
The riders earning maximum points for Stage `r stage_no`, and the teams that picked them were:
<br>

   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring riders for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
rider_stage_score <- tdf_table_master %>% 
  filter(stage == stage_no) %>% 
  select(rider, stage, "Total" = 12) %>% 
  distinct %>% 
  arrange(desc(Total))
# Display as neat table using 'kable' function
# No longer display this table, as I'm using the new table calculated below.
# kable(rider_stage_score[1:10, ])

```
<br>

``` {r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Next add a column with the names of teams against the high scoring stage riders:
for(r in 1:length(rider_stage_score$rider)){
  rider_stage_score[r, 4] <- tdf_table_master %>% 
    filter(rider == rider_stage_score$rider[r]) %>% 
    distinct(team) %>% 
    unlist() %>% 
    paste(collapse = ",")
}
# Give the new column a name
colnames(rider_stage_score)[4] <- "Selected by"

# Display the resulting table
kable(rider_stage_score[1:5, ])

```

<br>
Another take on rider data is provided in the chart below displaying Rider Cost vs the their Total Points accrued so far. The size of each circle is proportional the the number of teams that have selected that rider. 

   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# Top scoring riders cumulative

# Build a table with unique riders, times selected, total points and cost
  
  # First build a table with individual rider totals
  new_table <- tdf_table_master %>% 
    select(rider, Team,  stage, Cost, "Total" = 12) %>% 
    distinct() %>% 
    group_by(rider, Team, Cost) %>% 
    summarise("Total" = sum(Total)) %>% 
    arrange(rider)   # Arrange by rider so order is matched to next table
  # View(new_table)
  
  
  # Next build table with the number of times each rider has been selected
  rider_times_selected <- tdf_table_master %>% 
    filter(stage == 1) %>% 
    group_by("Rider" = rider) %>% 
    summarise("Times Selected" = n()) %>% 
    arrange(Rider)   # Again arrange by rider to match order from previous table
  # View(rider_times_selected)
  
  # Add times selected column from 2nd table to 1st table
  new_table[ , 5] <- rider_times_selected$`Times Selected`
  colnames(new_table)[5] <- "Times_Selected"
  
  # Create plot showing Total Points, Cost and Times Selected for each rider
  new_plot <- ggplot(data = new_table, aes(x = Cost, y = Total)) +
    geom_point(aes(size = Times_Selected, colour = "blue")) + 
    geom_text(data = subset(new_table, (Total > 100) | (Times_Selected >2)) ,   aes(label=rider),hjust=0, vjust=0, size = 4) + 
    theme(legend.position="bottom") + scale_size(range = c(2,12)) + 
    labs(title = "Rider Points vs Cost",   y = "Total Rider Points") 
  
  new_plot


```
   
<br>   
#### Top scoring teams for stage
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

# Top scoring teams for stage

# This dplyr subset looks at the total points for each
# rider for the stage and lists them in descending order.
team_stage_score <- tdf_table_master %>% 
  filter(stage == stage_no) %>% 
  select(team, directeur, stage, "Total" = 12) %>% 
  group_by(team, directeur) %>% 
  summarise("Stage_Points" = sum(Total)) %>% 
  arrange(desc(Stage_Points))

```
   
<br>  
So `r team_stage_score[1,2]` cleaned up stage `r stage_no`.
<br>  
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# Display as neat table using 'kable' function
kable(team_stage_score)
# Team scores by stage
team_scores <- tdf_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  arrange(stage, desc(`sum(Tot)`))
ggplot(data = team_scores, aes(x = stage, y = `sum(Tot)`)) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Stage Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

```
   
<br>  
The picture for overall after `r stage_no` stages is:
<br>
   
```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

### Top scoring teams cumulative

# Cumulative team scores by stage
team_cu_scores <- tdf_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  group_by(team) %>%
  mutate(cu_stage = cumsum(`sum(Tot)`))
# View(team_cu_scores)

# Plot cumulative team scores
ggplot(data = team_cu_scores, aes(x = stage, y = cu_stage, order = (team_cu_scores$team))) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Cumulative Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

team_overall <- team_cu_scores %>%
  filter(stage == stage_no) %>% 
  select("Team" = team, "Total" = cu_stage) %>% 
  arrange(desc(Total))
kable(team_overall)
# View(team_overall)

```
   
<br>
   
#### Inside Look
   
Here's a look inside the teams currently holding top spots on the GC podium. 

Actually - I can't get this bit of code to work for some frustrating reason, so I'm leaving it out and sending the update without it.
<br>  
   
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dlplyr filter looks at the top 4 scoring riders from
# the three teams that made the podium.

# This bit is crapping itself during the 'knit'. Needs work....
# stage_cu_podium <- team_cu_scores %>%
#   filter(stage == stage_no) %>%
#   arrange(desc(cu_stage))
# View(stage_cu_podium)

# team_cu_rider_scores <- tdf_table_master %>%
#   group_by(team, rider) %>%
#   summarise("Rider Total" = sum(Tot)) %>%
#   filter(team %in% stage_cu_podium$team[1:3]) %>%
#   arrange(team, desc(`Rider Total`)) %>%
#   top_n(n = 3, wt = `Rider Total`) %>%
#   rename("Rider" = rider, "Team" = team)
# View(team_cu_rider_scores)
# kable(team_cu_rider_scores, caption = "Rider Scores - Top Teams")

```

<br>
Next up is Stage `r (stage_no + 1)`.
<br>
`r paste("http://www.cyclingnews.com/tour-de-france/stage-", (stage_no + 1), "/", sep = "")`
<br>

### Team Cross Reference
<br>

Here's the team breakdown for each of squads. Start dissecting!
<br>


``` {r, echo=FALSE, warning=FALSE, message=FALSE}

  # Build a list of the riders selected by the mini-league and the number
  # of times they have been selected
  rider_list <- tdf_table_master %>% 
    filter(stage == 1) %>% 
    group_by(rider) %>% 
    summarise("Times_Selected" = n()) %>% 
    arrange(desc(Times_Selected))
  
  # Build a list of the directeurs in the mini-league
  directeur_list <- tdf_table_master %>% 
    select(directeur) %>% 
    distinct() %>% 
    arrange(directeur)
  
  # Now create the cross-reference matrix which shows which riders have been selected by which directeurs
  for (d in 1:length(directeur_list$directeur)){   # FOR loop to run through the list of team directeurs
    rider_list[ , d+2] <- NA
    colnames(rider_list)[(d+2)] <- toString(directeur_list$directeur[d])
    
    # Build list of team rider for specific team directeur 'd'
    team_list <- tdf_table_master %>% 
      filter(directeur == directeur_list$directeur[d]) %>% 
      select(rider) %>% 
      distinct()
    
    # Next, for each directeur, run through the complete list of mini-league rideres
    # and check to see if they are in 
    for (r in 1:length(rider_list$rider)){   # FOR loop to run through the total list of riders in the mini-league
      if(unlist(rider_list[r, 1]) %in% unlist(team_list)){
        rider_list[r, d+2] <- "Yes"
      } else  # End IF statement checking to see if rider exists in the directeur's team
      {rider_list[r, d+2] <- ""}
    }   # End FOR loop 'r'
  }   # END FOR loop running through all of the directeurs
  


kable(rider_list)

```






