---
title: "4051 Rouleurs 2016 TDF Recap"
date: "01 July 2017"
output: 
  html_document:
    theme: "cosmo"   # I also like 'spacelab' and 'yeti'
    highlight: "default"   # I also like 'tango'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

With the 2017 Tour De France set to commence wheels rolling tonight, it's perhaps a good time to reflect on our performance twelve months earlier. 

Yes - I came last..... by a long way..... and I'm still sending this out. 

With our 6 teams there was the opportunity for the selection of 54 unique riders in total. However the short list of favoured athletes in each category ultimately led us to select our squads from only 59 of the 198 riders starting the Tour at the Mont Saint-Michel island monastery on the 2nd of July, 2016.

```{r cars, echo=FALSE, warning=FALSE, message=FALSE}

# Open .csv file with Wilston Dads Giro 2016 data
setwd("/home/a_friend/data_analysis/projects/velogames/2016 TDF Recap/")
# CS - 712804410714, 4051 - 552723443913, AusBus - 911923153215
giro_table_master <- read.csv(file = "tdf_table_master_712804410714.csv", sep = ",")
colnames(giro_table_master) <- gsub("\\.", "", colnames(giro_table_master))

require(dplyr)
require(knitr)

# Calculate which riders were the most selected by our teams
rider_cost <- giro_table_master %>% 
  filter(stage == 1) %>% 
  group_by("Rider" = rider) %>% 
  summarise("Times Selected" = n()) %>% 
  arrange(desc(`Times Selected`))
kable(rider_cost[1:12,])

# View(giro_table_master)
```

The most popular rider was Columbian `r rider_cost$Rider[1]` with `r rider_cost[1,2]` of us selecting Nairoman to do battle on French roads.

### Top Ten Riders

But of course, picking riders on reputation and past performance was no guarantee of success in the first of the year's Grand Tours. The top ten highest scoring riders for the Giro, in order were:

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# require(dplyr)
rider_scores_dplyr <- giro_table_master %>% 
  select(rider, stage, Tot) %>% 
  distinct %>% 
  group_by(rider) %>% 
  summarise(sum(Tot)) %>% 
  arrange(desc(`sum(Tot)`))
# View(rider_scores_dplyr)

require(ggplot2)
ggplot(data = rider_scores_dplyr[1:10, ], aes(x = factor(rider_scores_dplyr$rider[1:10]), y = rider_scores_dplyr$`sum(Tot)`[1:10])) + geom_bar(stat = "identity", fill = "pink") + scale_x_discrete(limits = rider_scores_dplyr$rider[1:10]) + xlab("Rider") + ylab("Score") + theme(axis.text.x = element_text(angle = 60, hjust = 1))


```


The top ten highest scoring riders for the Giro, based on Return on Investment were:

``` {r, echo=FALSE, warning=FALSE, message=FALSE}

rider_value_dplyr <- giro_table_master %>% 
  mutate(ROI = Tot / Cost) %>% 
  select(rider, stage, ROI) %>% 
  distinct %>% 
  group_by(rider) %>% 
  summarise(sum(ROI)) %>% 
  arrange(desc(`sum(ROI)`))
# View(rider_value_dplyr)

ggplot(data = rider_value_dplyr[1:10, ], aes(x = factor(rider_value_dplyr$rider[1:10]), y = rider_value_dplyr$`sum(ROI)`[1:10])) + geom_bar(stat = "identity", fill = "blue") + scale_x_discrete(limits = rider_value_dplyr$rider[1:10]) + xlab("Rider") + ylab("ROI") + theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

The ROI data is insightful in that it reveals which riders provided the best bang for buck. Although expensive picks such as Froome ultimately proved good investments, the team backing Jarlinson Pantano must have been cheering all the way to the Velocredit bank!


### Overall Team Performance

At the end of a tightly fought three week battle the Bandits skipped away with it. However it required bonus point allocations on the final stage for Dick's team to seal victory.

``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dplyr subset looks at the total points for each
# team overall and lists them in descending order.
team_total <- giro_table_master %>% 
  group_by("Team" = team, "Directeur" = directeur) %>% 
  summarise(sum(Tot)) %>% 
  arrange(desc(`sum(Tot)`)) %>% 
  rename("Total" = `sum(Tot)`)
# Display as neat table using 'kable' function
kable(team_total)

```


### Individual Stage Performance



``` {r, echo=FALSE, warning=FALSE, message=FALSE}

# Team scores by stage
team_scores <- giro_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  arrange(stage, desc(`sum(Tot)`))
ggplot(data = team_scores, aes(x = stage, y = `sum(Tot)`)) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Stage Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

# Cumulative team scores by stage
team_cu_scores <- giro_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  group_by(team) %>%
  mutate(cu_stage = cumsum(`sum(Tot)`))
# Determine final placings
final_placings <- team_cu_scores %>%
  filter(stage == 21) %>%
  select(team, cu_stage) %>%
  arrange(desc(cu_stage)) %>%
  select(team)
# as.list(final_placings)

# View(team_cu_scores)
# Plot cumulative team scores
ggplot(data = team_cu_scores, aes(x = stage, y = cu_stage, order = (team_cu_scores$team))) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Cumulative Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

```

There was a real mix in rider selection in the top teams. 

``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dlplyr filter looks at the top 4 scoring riders from
# the three teams that made the podium.
team_cu_rider_scores <- giro_table_master %>% 
  group_by(team, rider) %>% 
  summarise("Rider Total" = sum(Tot)) %>% 
  filter(team %in% final_placings$team[1:3]) %>% 
  arrange(team, desc(`Rider Total`)) %>% 
  top_n(n = 4, wt = `Rider Total`) %>% 
  rename("Rider" = rider, "Team" = team)
kable(team_cu_rider_scores)

```


The bottom three teams all selected riders enjoying a successful Tour and associated large total points, but they just didn't have enough of them! Yes - my team of losers finished holding supporting the rest of the table......

``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dlplyr filter looks at the top 4 scoring riders from
# the three teams that made the podium.
team_bt_rider_scores <- giro_table_master %>% 
  group_by(team, rider) %>% 
  summarise("Rider Total" = sum(Tot)) %>% 
  filter(team %in% final_placings$team[(nrow(final_placings)-2):nrow(final_placings)]) %>% 
  arrange(team, desc(`Rider Total`)) %>% 
  top_n(n = 4, wt = `Rider Total`) %>% 
  rename("Rider" = rider, "Team" = team)
# Display as neat table using 'kable'
kable(team_bt_rider_scores)

````


Perhaps there's some lessons in there as you look to finalise your squads for this year's Tour. Get your team in before the cutoff early this evening!

