---
title: "AusBus 2016 Vuelta Recap"
date: "19 August 2017"
output: 
  html_document:
    theme: "cosmo"   # I also like 'spacelab' and 'yeti'
    highlight: "default"   # I also like 'tango'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

With the 2017 Vuelta a España set to commence wheels rolling tonight, it's perhaps a good time to reflect on our performance twelve months earlier. 

With 14 teams there was the opportunity for 9 unique teams and 81 riders in total. However the short list of favoured athletes in each category ultimately led us to select our squads from only 49 of the 198 riders starting the team time trial to Castrelo de Miño on the 20th of August, 2016.


```{r cars, echo=FALSE, warning=FALSE, message=FALSE}

# Open .csv file with Wilston Dads Giro 2016 data
setwd("/home/a_friend/data_analysis/projects/velogames/2016 Vuelta Recap/")
# CS - 131601094614, 4051 - 741602135413, AusBus - 231701090915
vuelta_table_master <- read.csv(file = "vuelta_table_master_231701090915.csv", sep = ",")
colnames(vuelta_table_master) <- gsub("\\.", "", colnames(vuelta_table_master))

require(dplyr)
require(knitr)

# Calculate which riders were the most selected by our teams
rider_cost <- vuelta_table_master %>% 
  filter(stage == 1) %>% 
  group_by("Rider" = rider) %>% 
  summarise("Times Selected" = n()) %>% 
  arrange(desc(`Times Selected`))
kable(rider_cost[1:12,])

# View(vuelta_table_master)
```

The most popular rider was Columbian `r rider_cost$Rider[1]` with `r rider_cost[1,2]` of us selecting the always smiling Orica team member to do battle on Spanish roads.

### Top Ten Riders

But of course, picking riders on reputation and past performance was no guarantee of success in the last of the year's Grand Tours. The top ten highest scoring riders for the Vuelta, in order were:

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# require(dplyr)
rider_scores_dplyr <- vuelta_table_master %>% 
  select(rider, stage, Tot) %>% 
  distinct %>% 
  group_by(rider) %>% 
  summarise(sum(Tot)) %>% 
  arrange(desc(`sum(Tot)`))
# View(rider_scores_dplyr)

require(ggplot2)
ggplot(data = rider_scores_dplyr[1:10, ], aes(x = factor(rider_scores_dplyr$rider[1:10]), y = rider_scores_dplyr$`sum(Tot)`[1:10])) + geom_bar(stat = "identity", fill = "pink") + scale_x_discrete(limits = rider_scores_dplyr$rider[1:10]) + xlab("Rider") + ylab("Score") + theme(axis.text.x = element_text(angle = 60, hjust = 1))


```

It's interesting to see that while Contador finished just off the podium in fourth place, he was a bit of a grey-man through the stages and didn't actually acrue a lot of Velocredits or even finish in the top ten riders based on ROI. Speaking of which...


The top ten highest scoring riders for the Vuelta, based on Return on Investment were:

``` {r, echo=FALSE, warning=FALSE, message=FALSE}

rider_value_dplyr <- vuelta_table_master %>% 
  mutate(ROI = Tot / Cost) %>% 
  select(rider, stage, ROI) %>% 
  distinct %>% 
  group_by(rider) %>% 
  summarise(sum(ROI)) %>% 
  arrange(desc(`sum(ROI)`))
# View(rider_value_dplyr)

ggplot(data = rider_value_dplyr[1:10, ], aes(x = factor(rider_value_dplyr$rider[1:10]), y = rider_value_dplyr$`sum(ROI)`[1:10])) + geom_bar(stat = "identity", fill = "blue") + scale_x_discrete(limits = rider_value_dplyr$rider[1:10]) + xlab("Rider") + ylab("ROI") + theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

The ROI data is insightful in that it reveals which riders provided the best bang for buck. Although expensive picks such as Quintana ultimately proved good investments, teams backing Ruben Fernandez must have been cheering all the way to the Velocredit bank!


### Overall Team Performance

At the end of a tightly fought three week battle Steve's safety first helmets took top step on the podium by a comfortable margin over Billy's squad. 

``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dplyr subset looks at the total points for each
# team overall and lists them in descending order.
team_total <- vuelta_table_master %>% 
  group_by("Team" = team, "Directeur" = directeur) %>% 
  summarise(sum(Tot)) %>% 
  arrange(desc(`sum(Tot)`)) %>% 
  rename("Total" = `sum(Tot)`)
# Display as neat table using 'kable' function
kable(team_total)

```


### Individual Stage Performance

Steve's team emerged in a controlling fashion from about Stage 7 and didn't look like relinquishing the lead for the remainder of the race.  


``` {r, echo=FALSE, warning=FALSE, message=FALSE}

# Team scores by stage
team_scores <- vuelta_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  arrange(stage, desc(`sum(Tot)`))
ggplot(data = team_scores, aes(x = stage, y = `sum(Tot)`)) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Stage Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

# Cumulative team scores by stage
team_cu_scores <- vuelta_table_master %>% 
  group_by(stage, team) %>% 
  summarise(sum(Tot)) %>% 
  group_by(team) %>%
  mutate(cu_stage = cumsum(`sum(Tot)`))
# Determine final placings
final_placings <- team_cu_scores %>%
  filter(stage == 21) %>%
  select(team, cu_stage) %>%
  arrange(desc(cu_stage)) %>%
  select(team)
# as.list(final_placings)

# View(team_cu_scores)
# Plot cumulative team scores
ggplot(data = team_cu_scores, aes(x = stage, y = cu_stage, order = (team_cu_scores$team))) + geom_line(aes(colour = team), size = 1.2) + labs(title ="Cumulative Score", x = "Stage", y = "Points") + theme(legend.position = "bottom")

```

 
Steve's selection of big scoring Quintana and Valverde were differentiators for his squad.


``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dlplyr filter looks at the top 4 scoring riders from
# the three teams that made the podium.
team_cu_rider_scores <- vuelta_table_master %>% 
  group_by(team, rider) %>% 
  summarise("Rider Total" = sum(Tot)) %>% 
  filter(team %in% final_placings$team[1:3]) %>% 
  arrange(team, desc(`Rider Total`)) %>% 
  top_n(n = 4, wt = `Rider Total`) %>% 
  rename("Rider" = rider, "Team" = team)
kable(team_cu_rider_scores)

```

Dave Hazzard was reliant on Chaves to do his heavy lifting, but the points haul per rider fell away quickly after that. He saved Bail's sorry bunch from the bottom spot.


``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# This dlplyr filter looks at the top 4 scoring riders from
# the three teams that made the podium.
team_bt_rider_scores <- vuelta_table_master %>% 
  group_by(team, rider) %>% 
  summarise("Rider Total" = sum(Tot)) %>% 
  filter(team %in% final_placings$team[(nrow(final_placings)-2):nrow(final_placings)]) %>% 
  arrange(team, desc(`Rider Total`)) %>% 
  top_n(n = 4, wt = `Rider Total`) %>% 
  rename("Rider" = rider, "Team" = team)
# Display as neat table using 'kable'
kable(team_bt_rider_scores)

````


Perhaps there's some lessons in there as you look to finalise your squads for this year's Vuelta Get your team in before the cutoff early this evening!

